---
title: 'Exploratory Data Analysis of NYC Flights 2023: Answering 10 Key Questions'
author: "Kanyaphat Lokhan"
date: "2025-06-22"
output:
  pdf_document: default
  html_document: default
---

## Install Package

First of all, we need to install the relevant packages, including nycflights23, which contains the flights dataset.
```{r,message=FALSE}
install.packages('dplyr')
install.packages('tidyverse')
install.packages('nycflights23')

library(dplyr)
library(tidyverse)
library(nycflights23)
```

The flights contains information about all flights that departed from one of the three New York City airports (EWR, JFK, and LGA) in 2023. 
Let's have a look at the structure.

## Data overview
```{r}
str(flights)
```

## Column description
- **year, month, day**: Date of departure.

- **dep_time, arr_time**: Actual departure and arrival times, UTC.

- **sched_dep_time, sched_arr_time**: Scheduled departure and arrival times, UTC.

- **dep_delay, arr_delay**: Departure and arrival delays, in minutes. Negative times represent early departures/arrivals.

- **hour, minute**: Time of scheduled departure broken into hour and minutes.

- **carrier**: Two letter carrier abbreviation.

- **tailnum**: Plane tail number.

- **flight**: Flight number.

- **origin, dest**: Origin and destination airport. See get_airports for additional metadata.

- **air_time**: Amount of time spent in the air, in minutes.

- **distance**: Distance between airports, in miles.

- **time_hour**: Scheduled date and hour of the flight as a POSIXct date. Along with origin, can be used to join flights data to weather data.

## Key Questions
1. What is the average departure delay for flights originating from EWR (Newark Liberty International Airport) in the month of July?

```{r}
avg_dep_delay <- flights %>%
  filter(origin == "EWR" & month == 7 & !is.na(dep_delay) & dep_delay > 0) %>% 
  summarise(mean(dep_delay)) 

print(avg_dep_delay)
```

**Ans** The average delay for flights originating from EWR in July is **63.2** mins

2. How many flights were cancelled by each airline (carrier)? Show the carrier code and the count of cancelled flights.

```{r}
cancel_flights <- flights %>%
  filter(is.na(air_time) & is.na(dep_delay) & is.na(arr_delay)
         & is.na(dep_time) & is.na(arr_time)) %>%
	group_by(carrier) %>%
  summarise(cancelled_flights = n()) %>%
  arrange (desc(cancelled_flights))

print(cancel_flights)
```

**Ans** There are **14** airlines that have cancelled flights. Among them, YX airline has cancelled the most flights, with 3,062 cancellations, while HA airline has cancelled the fewest, with only 2 flights.

3. Which destination airport (dest) had the highest average arrival delay in the dataset?

```{r}
highest_avg_arr_delay <- flights %>%
  select(dest, arr_delay) %>%
  filter (arr_delay > 0, !is.na(arr_delay)) %>%
  group_by (dest) %>%
  summarise(avg_arr_delay = mean(arr_delay)) %>%
  arrange (desc(avg_arr_delay)) %>%
  head(1)

print(highest_avg_arr_delay)
```
**Ans** **EGE** had the highest average arrival delay, which is 74.9 mins.

4. Create a new column called total_delay which is the sum of dep_delay and arr_delay. What is the median total_delay for flights in the dataset?

```{r}
med_delay <- flights %>%
  mutate(total_delay = dep_delay + arr_delay) %>%
  filter(!is.na(total_delay)) %>%
  summarise(median_delay = median(total_delay)) 

print(med_delay)
```

**Ans** Median total delay for flights is **-12**

5. For flights departing in the first quarter of the year (January-March), what percentage of flights departed on time (dep_delay <= 0)?

```{r}
Q1_on_time <- flights %>%
  filter (!is.na(dep_delay) & month %in% 1:3) %>%
  summarise (total_flight = n(),
             total_on_time_flight = sum(dep_delay <= 0),
             dep_on_time_firstQ = (total_on_time_flight/total_flight)*100)

print(Q1_on_time)
```

**Ans** The percentage of flights departed on time in the first quarter of 2023 is **63 %**

6. Which month had the highest number of flights departing from JFK (John F. Kennedy International Airport)?

```{r}
highest_dep <- flights %>%
  filter(origin == "JFK") %>%
  group_by(month) %>%
  summarise(number_of_flight = n()) %>%
  arrange(desc(number_of_flight)) %>%
  head(1)

print(highest_dep)
```

**Ans** **3rd month** had 12,158 flights departing from JFK, which is the highest number from 2023.

7. What is the longest distance flown by any flight in the dataset, and which flight (tailnum) achieved this?

```{r}
longest_flights <- flights %>%
  select(tailnum, origin, dest, distance) %>%
  filter(distance == max(distance, na.rm = TRUE))

print(longest_flights)
```

**Ans** The longest distance is **4,983 miles**.

8. Group the flights by carrier and calculate the average air time (air_time) for each carrier. Display the carrier and their average air time, sorted in descending order of average air time.

```{r}
avg_air_time <- flights %>%
  select(carrier, air_time) %>%
  group_by(carrier) %>%
  summarise (avg_air_time = mean(air_time, na.rm = TRUE)) %>%
  arrange(desc(avg_air_time))

print(avg_air_time)
```

9. How many flights departed before 6 AM (hour < 6)?

```{r}
dep_before_6 <- flights %>%
  select (carrier, hour, minute) %>%
  filter (hour < 6) %>%
  summarise(num_flights = n()) 

print(dep_before_6)
```

**Ans** There are **3,270 flights** departed before 6 AM.

10. Create a table showing the number of flights from each origin to each destination. Which origin-destination pair had the most flights?

```{r}
pair_flights <- flights %>%
  select (origin, dest) %>%
  group_by(origin, dest) %>%
  summarise(num_of_origin_dest_pair = n()) %>%
  arrange(desc(num_of_origin_dest_pair))

print(pair_flights)
```

**Ans** **JFK-LAX** had the most flights, which is 10045 flights.


